---
title: "Thesis"
author: "Miguel Reyes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# National Health and Nutrition Examination Survey

```{r libraries,echo= TRUE}
library(nhanesA)
library(readr)
library(predtools)
library(magrittr)
library(dplyr)
library(ggnewscale)
#library(caret)
library(InformationValue)
library(ISLR)
library(cvms)
library(tibble)
library(tidyverse)
library(rstatix)
library(ggpubr)
# install.packages("remotes")
# remotes::install_github("ML4LHS/runway")
library(runway)
library(ggmosaic)

setwd("~/Semester 2/Research project/Db")
```

## Demo

### Demographic Variables & Sample Weights
*	RIAGENDR – Gender
*	RIDAGEYR - Age at Screening Adjudicated – Recode
*	RIDAGEMN - Age in Months
*	RIDRETH2 - Linked NH3 Race/Ethnicity - Recode
*	DMDEDUC - Education - Recode (old version)
*	INDHHINC - Annual Household Income

```{r}
#nhanesTableVars('DEMO', 'DEMO')
demo  <- nhanes('DEMO')
select_cols <- c('SEQN', 'RIAGENDR', 'RIDAGEYR', 'RIDRETH2', 'DMDEDUC', 'INDHHINC')
demo <- demo[,select_cols]
colSums(is.na(demo))
demo <- demo[complete.cases(demo),]
```


### Body Measures (BMX)

•	BMXWT - Weight (kg)
•	BMXRECUM - Recumbent Length (cm)
•	BMXHEAD - Head Circumference (cm)
•	BMXHT - Standing Height (cm)
•	BMXBMI - Body Mass Index (kg/m**2)
•	BMXLEG - Upper Leg Length (cm)
•	BMXCALF - Maximal Calf Circumference (cm)
•	BMXARML - Upper Arm Length (cm)
•	BMXARMC - Arm Circumference (cm)
•	BMXWAIST - Waist Circumference (cm)
•	BMXTHICR - Thigh Circumference (cm)
•	BMXTRI - Triceps Skinfold (mm)
•	BMXSUB - Subscapular Skinfold (mm)

```{r}
#### Examination Data ####
# Body Measures (BMX)
#nhanesTableVars('EXAM', 'BMX')
bmx  <- nhanes('BMX')
select_cols <- c('SEQN', 'BMXWT', 'BMXHT', 'BMXBMI', 'BMXLEG'
                 , 'BMXCALF', 'BMXARML', 'BMXARMC', 'BMXWAIST', 'BMXTHICR', 'BMXTRI'
                 , 'BMXSUB')
bmx <- bmx[,select_cols]
bmx <- bmx[complete.cases(bmx),]
colSums(is.na(bmx))
```


## Laboratory Data

### Albumin & Creatinine – Urine (LAB16)
* All

```{r}
lab16  <- nhanes('LAB16')
select_cols <- c('SEQN', 'URXUMA')
lab16 <- lab16[,select_cols]
colSums(is.na(lab16))
lab16 <- lab16[complete.cases(lab16),]
```

### Blood Pressure (BPX)

•	PEASCTM1 - Blood Pressure Time in Seconds
•	BPQ150A - Have you had any of the following in the past 30 minutes?: Food
•	BPQ150B - Had alcohol in the past 30 minutes?
•	BPQ150C - Had coffee in the past 30 minutes?
•	BPQ150D - Had cigarettes in the past 30 minutes?
•	BPAARM - Arm selected
•	BPACSZ - Coded cuff size
•	BPXPLS - 60 sec. pulse (30 sec. pulse * 2)
•	BPXPULS - Pulse regular or irregular?
•	BPXML1 - MIL: maximum inflation levels (mm Hg)
•	BPXSAR - SBP average reported to examinee
•	BPXDAR - DBP average reported to examinee

```{r}
bpx  <- nhanes('BPX')
select_cols <- c('SEQN', 'PEASCTM1', 'BPXPLS','BPXML1','BPXSAR','BPXDAR')
bpx <- bpx[,select_cols]
colSums(is.na(bpx))
bpx <- bpx[complete.cases(bpx),]
```


### Glycohemoglobin (LAB10)
* All
```{r}
lab10  <- nhanes('LAB10')
colSums(is.na(lab10))
lab10 <- lab10[complete.cases(lab10),]
```



```{r}
################
#NHANES VERSION#
################

setwd("~/Semester 2/Research project/Db")
srvyin <- paste("NHANES_1999_2000_MORT_2019_PUBLIC.dat")   # full .DAT name here
srvyout <- "NHANES_1999" # shorthand dataset name here

# Example syntax:
#srvyin <- paste("NHANES_1999_2000_MORT_2019_PUBLIC.dat")   
#srvyout <- "NHANES_1999_2000"      

# read in the fixed-width format ASCII file
dsn <- read_fwf(file=srvyin,
                col_types = "iiiiiiii",
                fwf_cols(seqn = c(1,6),
                         eligstat = c(15,15),
                         mortstat = c(16,16),
                         ucod_leading = c(17,19),
                         diabetes = c(20,20),
                         hyperten = c(21,21),
                         permth_int = c(43,45),
                         permth_exm = c(46,48)
                ),
                na = c("", ".")
)


assign(paste0(srvyout), dsn)
rm(dsn, srvyin, srvyout)
```


```{r}
NHANES_1999<- NHANES_1999 %>% 
       mutate(d_months = case_when((NHANES_1999$eligstat ==1 & NHANES_1999$permth_exm < 120 ) ~ 1,
                                   (NHANES_1999$eligstat ==1 & NHANES_1999$permth_exm == NA ) ~ 100,
                                              TRUE ~ 0)) 
NHANES_1999 <- NHANES_1999 %>%
  naniar::replace_with_na(replace = list(x = c(100)))

NHANES_1999$mortstat <- NHANES_1999$d_months
```

```{r}
# Mortality
select_cols <- c('seqn', 'mortstat')
NHANES_1999 <- NHANES_1999[,select_cols]
colSums(is.na(NHANES_1999))
NHANES_1999 <- NHANES_1999[complete.cases(NHANES_1999),]
```


# Mergin data

```{r}

# Body Measures (BMX)
merger <- merge(demo, bmx)

merger<-merge(NHANES_1999, merger, by.x = 'seqn' , by.y = 'SEQN', all = TRUE)

# Albumin & Creatinine •	URXUMA      kidney disease
merger<-merge(merger, lab16, by.x = 'seqn' , by.y = 'SEQN', all = FALSE)

# Blood Pressure (BPX)
merger<-merge(merger, bpx, by.x = 'seqn' , by.y = 'SEQN', all = FALSE)

# Glycohemoglobin (lab10)
merger<-merge(merger, lab10, by.x = 'seqn' , by.y = 'SEQN', all = FALSE)


#defined before

merger <- merger[complete.cases(merger),]

merger$seqn=NULL
colSums(is.na(merger))
merger$RIDRETH2 <- as.factor(merger$RIDRETH2)
merger$RIAGENDR <- as.factor(merger$RIAGENDR)
head(merger)
```

```{r}
# Color by groups
library(ggplot2)
library(ggmosaic)
select_cols <- c('mortstat','RIDAGEYR','DMDEDUC','BPXDAR','BPXSAR','BMXARMC','BMXWAIST','RIAGENDR','URXUMA','BPXPLS')
descr_data <- merger[,select_cols]
descr_data$mortstat<-as.factor(descr_data$mortstat)
descr_data$RIAGENDR<- recode(descr_data$RIAGENDR, '1' = "Male",'2' = "Female")
# p1 <- ggplot2::ggplot(data = mydata, aes(x = intvar1, y = intvar2)) + geom_point() + ggtitle("Graph title")
# RIAGENDR
descr_data$DMDEDUC<-as.factor(descr_data$DMDEDUC)
ggplot(data = descr_data) +
  geom_mosaic(aes(x = product(RIAGENDR,y=mortstat), fill=RIAGENDR)) +  labs(title='Gender',x='Dead within 10 years',y='Gender') + scale_fill_manual(values = c('#56B4E9', "#FF9999")) +guides(fill=guide_legend(title="Gender"))+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

# RIDAGEYR
p1 <- ggplot2::ggplot(data = descr_data, aes(x = mortstat, y = RIDAGEYR, group=mortstat,color=mortstat)) + geom_boxplot() + ggtitle("Age") +  labs(title='RIDAGEYR = Age',x='Dead within 10 years',y='Age')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(color=guide_legend(title="Dead within 10 years"))
p2 <- ggplot(descr_data, aes(x = RIDAGEYR, fill = mortstat)) + geom_histogram() + ggtitle("Age") +  labs(title='RIDAGEYR = Age',x='',y='Age')+ theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(fill=guide_legend(title="Dead within 10 years"))
p1
p2
# BPXDAR
p1 <- ggplot2::ggplot(data = descr_data, aes(x = mortstat, y = BPXDAR, group=mortstat,color=mortstat)) + geom_boxplot() +
  labs(title='BPXDAR = SBD average reported',x='Dead within 10 years',y='SBD')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(color=guide_legend(title="Dead within 10 years"))
p2 <- ggplot(descr_data, aes(x = BPXDAR, fill = mortstat)) + geom_histogram() +labs(title='BPXDAR = SBD average reported',x='',y='SBD')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(fill=guide_legend(title="Dead within 10 years"))
p1
p2
# BPXSAR
p1 <- ggplot2::ggplot(data = descr_data, aes(x = mortstat, y = BPXSAR, group=mortstat,color=mortstat)) + geom_boxplot() + labs(title='BPXSAR = SBP average reported',x='Dead within 10 years',y='SBP')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(color=guide_legend(title="Dead within 10 years"))
p2 <- ggplot(descr_data, aes(x = BPXSAR, fill = mortstat)) + geom_histogram()  + labs(title='BPXSAR = SBP average reported',x='',y='SBP')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(fill=guide_legend(title="Dead within 10 years"))
p1
p2
# BMXARMC
p1 <- ggplot2::ggplot(data = descr_data, aes(x = mortstat, y = BMXARMC, group=mortstat,color=mortstat)) + geom_boxplot() +  labs(title="BMXARMC = Arm Circumference",x='Dead within 10 years',y='Arm Circumference')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(color=guide_legend(title="Dead within 10 years"))
p2 <- ggplot(descr_data, aes(x = BMXARMC, fill = mortstat)) + geom_histogram() +  labs(title='BMXARMC = Arm Circumference',x='',y='Arm Circumference')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(fill=guide_legend(title="Dead within 10 years"))
p1
p2
# BMXWAIST
p1 <- ggplot2::ggplot(data = descr_data, aes(x = mortstat, y = BMXWAIST, group=mortstat,color=mortstat)) + geom_boxplot() +  labs(title="BMXWAIST = Waist Circumference",x='Dead within 10 years',y='Waist Circumference')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(color=guide_legend(title="Dead within 10 years"))
p2 <- ggplot(descr_data, aes(x = BMXWAIST, fill = mortstat)) + geom_histogram() +  labs(title="BMXWAIST = Waist Circumference",x='',y='Waist Circumference')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(fill=guide_legend(title="Dead within 10 years"))
p1
p2
# URXUMA
p1 <- ggplot2::ggplot(data = descr_data, aes(x = mortstat, y = URXUMA, group=mortstat,color=mortstat)) + geom_boxplot() +  labs(title="URXUMA = Albumin & Creatinine",x='Dead within 10 years',y='Albumin & Creatinine')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(color=guide_legend(title="Dead within 10 years"))+ ylim(0, 100)
p2 <- ggplot(descr_data, aes(x = URXUMA, fill = mortstat)) + geom_histogram()+ xlim(0, 100) +  labs(title="URXUMA = Albumin & Creatinine",x='',y='Albumin & Creatinine')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(fill=guide_legend(title="Dead within 10 years"))
p1
p2

# BPXPLS
p1 <- ggplot2::ggplot(data = descr_data, aes(x = mortstat, y = BPXPLS, group=mortstat,color=mortstat)) + geom_boxplot() +  labs(title="BPXPLS =60 sec. ",x='Dead within 10 years',y='60 sec. pulse')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(color=guide_legend(title="Dead within 10 years")) + ylim(0, 100)
p2 <- ggplot(descr_data, aes(x = BPXPLS, fill = mortstat)) + geom_histogram() + labs(title="BPXPLS =60 sec. pulse ",x='',y='60 sec. pulse')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16) ) +guides(fill=guide_legend(title="Dead within 10 years"))+xlim(0, 100)
p1
p2
```



# First attempt 

```{r}
# Loading package
library(caTools)
library(ROCR) 
set.seed(12345)

# Splitting dataset
split <- sample.split(merger, SplitRatio = 0.5)

   
train_reg <- subset(merger, split == "TRUE")
test_reg <- subset(merger, split == "FALSE")
   
# Training model
logistic_model <- glm(mortstat ~ .,
                      data = train_reg,
                      family = "binomial")

intercept_only <- glm(mortstat ~ 1, data=train_reg,
                      family = "binomial")
all <- glm(mortstat ~ ., data=train_reg,
                      family = "binomial")
logistic_model <- step(intercept_only, direction='forward', scope=formula(all), trace=0)

# Summary
summary(logistic_model)
   
# Predict test data based on model
predict_reg <- predict(logistic_model, 
                       test_reg, type = "response")

# Changing probabilities
optimal <- optimalCutoff(test_reg$mortstat, predict_reg, optimiseFor = "misclasserror")[1]

predict_reg <- ifelse(predict_reg >optimal, 1, 0)
print(optimal)
# Evaluating model accuracy
# using confusion matrix
table(test_reg$mortstat, predict_reg)
missing_classerr <- mean(predict_reg != test_reg$mortstat)
tidy(logistic_model)
```



```{r}
tab_matrix<-table(tibble("Actual" = test_reg$mortstat,
       "prediction" = predict_reg))
cfm <- as_tibble(tab_matrix)
plot_confusion_matrix(cfm, 
                      target_col = "Actual", 
                      prediction_col = "prediction",
                      counts_col = "n")
```

```{r}
library(randomForest)
library(datasets)
#library(caret)
#install.packages('datasets')
model1 <- randomForest(mortstat ~ ., data = train_reg, importance = TRUE,ntree=500)

predTest <- predict(model1, test_reg, type = "response")
optimal <- optimalCutoff(test_reg$mortstat, predTest,optimiseFor = "misclasserror")[1]
optimal
predTest <- ifelse(predTest >optimal, 1, 0)
tab_matrix<-table(tibble("prediction" = predTest,"Actual" = test_reg$mortstat))
cfm <- as_tibble(tab_matrix)
plot_confusion_matrix(cfm, 
                      target_col = "Actual", 
                      prediction_col = "prediction",
                      counts_col = "n")
```

```{r}

plot(model1, col='red',main="Random Forest")
```



```{r}
test_reg$Regression   <- predict(logistic_model, test_reg, type = "response")
calibration_plot(data = test_reg, obs = "mortstat", pred = "Regression", title = "Calibration plot for Regression model")

test_reg$RandomForest <- predict(model1, test_reg, type = "response")
calibration_plot(data = test_reg, obs = "mortstat", pred = "RandomForest", title = "Calibration plot for RandomForest model")
```

```{r}

cal_plot(test_reg, outcome='mortstat',prediction='Regression')
cal_plot(test_reg, outcome='mortstat',prediction='RandomForest')

```

```{r}
library(ROCR)
library(pROC)
pred <- prediction(test_reg$Regression, test_reg$mortstat)
perf <- performance(pred,"tpr","fpr")
#plot(perf,colorize=TRUE, title = "plot for regression model")

pred <- prediction(test_reg$RandomForest, test_reg$mortstat)
perf2 <- performance(pred,"tpr","fpr")
#plot(perf2,colorize=TRUE, title = "plot for RandomForest model")

rocobj <- roc(test_reg$mortstat, test_reg$Regression)
ggroc(rocobj)

plot(perf, col='red', xlab="1-Specificity", ylab="Sensitivity")
plot(perf2, add = TRUE, col='green')
legend("bottomright", 95, legend=c("Logistic Regression", "Random Forest"),
       col=c("red", "green"), lty=1, cex=1)

pred <- prediction(test_reg$Regression, test_reg$mortstat)
auc.tmp<-performance(pred,"auc")
auc_regression <- as.numeric(auc.tmp@y.values)
pred <- prediction(test_reg$RandomForest, test_reg$mortstat)
auc.tmp<-performance(pred,"auc")
auc_raforest <- as.numeric(auc.tmp@y.values)
  
auc_regression
auc_raforest
```
#### SIMULATION ####

```{r warning=FALSE}
# Taking the sample
set.seed(145)
auc_reg=vector()
auc_for=vector()

n_sim=500

for (i in 1:n_sim){
  sample_nhanes <-sample_n(merger, 2000, replace = FALSE)

  # Split train and test
  split <- sample.split(sample_nhanes, SplitRatio = 0.7)
  train <- subset(sample_nhanes, split == "TRUE")
  test <- subset(sample_nhanes, split == "FALSE")
  
  # stepwise regression
  intercept_only <- glm(mortstat ~ 1, data=train,
                      family = "binomial")
  all <- glm(mortstat ~ ., data=train,
                      family = "binomial")
  regression_model <- step(intercept_only, direction='forward', scope=formula(all), trace=0)
  
  # Random forest
  forest_model <- randomForest(mortstat ~ ., data = train, importance = TRUE)
  
  #auc
  test$Regression <- predict(regression_model, test, type = "response")
  test$Forest <- predict(forest_model, test, type = "response")
  
  
  pred <- prediction(test$Regression, test$mortstat)
  auc.tmp<-performance(pred,"auc")
  auc_regression <- as.numeric(auc.tmp@y.values)
  auc_reg[i] <- auc_regression
  
  pred <- prediction(test$Forest, test$mortstat)
  auc.tmp<-performance(pred,"auc")
  auc_forest <- as.numeric(auc.tmp@y.values)
  auc_for[i]<-auc_forest
}

mean(auc_reg)
mean(auc_for)

```

```{r}
cum_auc_reg <- cumsum(auc_reg) / seq_along(auc_reg)
cum_auc_for <- cumsum(auc_for) / seq_along(auc_for)


plot(1:500, cum_auc_reg,
main="AUC average",
ylab="AUC average",
xlab="number of simulations",
ylim = c(0.78,1),
type="l",
col="red")
lines(1:500,cum_auc_for, col="blue")
legend("topright",
c("Logistic Regression","Random Forest"),
fill=c("red","blue")
)

```
```{r}
sample_size<-c(250,500,750,1000,1250,1500,1750,2000)
auc_r<-c(0.8349055,0.9055964,0.9118371,0.9179247,0.9165847,0.9190276,0.9211192,0.9215774)
auc_f<-c(0.8863919,0.9054373,0.9098305,0.9120371,0.9133123,0.9145869,0.9172978,0.9168623)

performance_r_f<-data.frame(sample_size,auc_r,auc_f)

knitr::kable(performance_r_f)
```


```{r}
linear_reg <- lm(auc_r ~ sample_size, data=performance_r_f)
knitr::kable(tidy(linear_reg))
linear_for <- lm(auc_f ~ sample_size, data=performance_r_f)
knitr::kable(tidy(linear_for))
```
```{r}
p <- ggplot(performance_r_f,aes(sample_size, auc_r,col='Regression')) +
  geom_smooth(method='lm', formula= y~x) +
  new_scale_color() +
  geom_smooth(aes(sample_size, auc_f,col='Random Forest'),method='lm', formula= y~x,alpha=.3) +
  scale_color_manual(values = c("blue"))+
  labs(title = "", x = "Sample size", y = "AUC")
p
```
```{r}
p <- ggplot(performance_r_f,aes(sample_size, auc_r,col='Regression')) + geom_point() +
  new_scale_color() + geom_point(aes(sample_size, auc_f,col='Random Forest')) +
  scale_color_manual(values = c("blue"))+
  labs(title = "", x = "Sample size", y = "AUC")
p
```

